An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-23 neurips-19/poly-embed.tex
comment or two in sec2
Iteration over top k proof in Appendix

We have now seen the tight relationship between polyhedral losses and embeddings; in particular, every polyhedral loss embeds some discrete loss.
The embedding itself tells us how to map the embedded points back to the discrete reports (map $\varphi(r)$ to $r$), but it is not clear when this map can be extended to the remaining reports to obtain a full link function, and whether such a link can lead to consistency.
In this section, we give a construction to generate calibrated links for \emph{any} polyhedral loss.

The first step is to give a link $\psi$ such that exactly minimizing expected surrogate loss $L$, followed by applying $\psi$, always exactly minimizes expected original loss $\ell$.
\bo{Maybe cite prop in the appendix}
The existence of such a link is somewhat subtle, because in general some point $u$ that is far from any embedding point can minimize expected loss for two very different distributions $p,p'$, making it unclear whether there exists a link choice $\psi(u)$ that is simultaneously optimal for both.
We show\bo{cite lemma} that as we vary $p$ over the infinite $\simplex$, there are only finitely many sets of the form $U = \argmin_{u \in \reals^d} \langle p, L(u) \rangle$.
Associate each $U$ with $R_U \subseteq \R$, the set of reports whose embedding points are in $U$.
Now enforce that all points in $U$ link to some report in $R_U$.
(As a special case, embedding points must link to their corresponding reports.)
Proving this is well-defined uses a chain of arguments involving the Bayes risk\bo{characterization prop in appendix, ...}, ultimately showing that if $u$ lies in multiple $U$, the corresponding report sets $R_U$ all intersect at some $r =: \psi(u)$.

%
%\begin{proposition}\label{prop:embed-link}
%  Let $L$ elicit $\Gamma:\simplex\toto\reals^d$ which embeds a finite property $\gamma$.
%  Then there exists a link function $\psi: \reals^d \to \R$ that is \emph{valid} in the sense that, for any $p$, if $u$ minimizes expected $L$-loss (i.e. $u \in \Gamma(p)$, then $\psi(u)$ minimizes expected $\ell$-loss (i.e. $\psi(u) \in \gamma(p)$).
%\end{proposition}
%\begin{proof}
%  Let $\gamma:\simplex\toto\R$.
%  Proposition~\ref{prop:embed-trim} gives us that $\trim(\Gamma) = \{\gamma_r : r\in\R\}$.
%  \bo{Have we defined trim already? Or is this proof going to the appendix?}
%  Then for any $u\in\reals^d$, there is some $r\in\R$ such that $\Gamma_u \subseteq \gamma_r$.
%  The link $\psi : \reals^d \to \R$ which encodes these choices is valid.
%\end{proof}

Intuitively, to create separation, we just need to ``thicken'' this construction by mapping all approximately-optimal points $u$ to optimal reports $r$.
Let $\U$ contain all optimal report sets $U$ of the form above.

%\begin{lemma}
%  \label{lem:polyhedral-range-gamma}
%  Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss, and let $\Gamma = \prop{L}$.
%  Then the range of $\Gamma$, $\U = \Gamma(\simplex) = \{\Gamma(p) \subseteq \reals^d | p\in\simplex\}$, is a finite set of closed polyhedra.
%\end{lemma}
%For each of these finitely many $U \in \U$, we let $R_U$ be the set of reports whose embedding points lie in $U$.
%The idea will be to map points in $U$, as well as points nearby, to something in $R_U$
%We show in the full version \bo{okay wording?} that for any $p$, if $U = \Gamma(p)$, then $R_U = \gamma(p)$.

@@

\begin{proof}[Sketch]
  \emph{Well-defined:}
  For the initial construction above, we argued that if some collection such as $U,U',U''$ overlap at a $u$, then their report sets $R_U, R_{U'}, R_{U''}$ also overlap, so there is a valid choice $r = \psi(u)$.
  Now, we thicken all sets $U \in \U$ by a small enough $\epsilon$; it can be shown that if the \emph{thickened} sets overlap at $u$, then $U,U',U''$ themselves overlap, so again $R_U, R_{U'}, R_{U''}$ overlap and there is a valid chioce $r = \psi(u)$.

  \emph{Calibrated:} By construction of the thickened link, if $u$ maps to a wrong report, i.e. $\psi(u) \not\in \gamma(p)$, then $u$ must have at least distance $\epsilon$ to the optimal set $U$.
  We then show that the minimal gradient of the expected loss along any direction away from $U$ is lower-bounded, giving a constant excess expected loss at $u$.
\end{proof}

@@

The top-$k$ classification problem takes as input a set $r \in \R := \{r \subseteq [n] : |r| = k\}$, and observes outcome $y \in [n]$.
We define the top-$k$ loss $\ell_{top-k}:\R \to \reals$ by
\begin{align}\label{eq:top-k}
\ell_{top-k}(r,y) = \begin{cases}
0 & y \in r\\
1 & \text{otherwise}
\end{cases}
\end{align}

\citet{lapin2016loss, lapin2018analysis}, provide two surrogate losses for measuring top-$k$ error, one of which is shown to be convex, but their consistency is left as an open question.
\citet{yang2018consistency} show the inconsistency of their losses, and create a new surrogate loss that is consistent for the top-$k$ problem, but it should be noted that Yang and Koyejo's surrogate is not convex.
They also show that Lapin's surrogates are not consistent, but without too much intuition as to why.
Following the notation of Yang, we call Lapin's~\cite{lapin2018analysis} convex surrogate $L_2$, and focus on this.
Here, we assume $r \subseteq [n]$ and $k \leq n$ is a constant.

We show that $L_2$ embeds some other discrete loss $\ell_2:\{0,1\}^n \to \reals^\Y_+$, which has a different Bayes Risk than $\ell_{0-1}$.

In order to do this, we want to show that for all $p \in \simplex$, $\inf_{u' \in \reals^n} \inprod{p}{L_2(u')} = \min_{u' \in \{0,1\}^n} \inprod{p}{L_2(u')}$. (Lemma~\ref{lem:top-k-optimal-corners}.)
Then, we show there is a discrete loss $\ell_2 : \{0,1\}^n \to \reals^\Y_+$ that is embedded by $L_2$.  (Lemma~\ref{lem:top-k-surrogate-embeds}.)

We conclude with a concrete example; $k=1$.
One can see that the $\ell_{top-1}$ reduces to 0-1 loss, and we characterize the distributions $p \in \simplex$ where $\risk{L_2}(p) = \risk{\ell_2}(p) \neq \risk{\ell_{top-k}}(p)$.

@@

We are given the surrogate 
\begin{align}\label{eq:L-2-surrogate}
L_2(u,y) := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+
\end{align}
where $x_{[i]}$ is the $i^{th}$ largest component of $x$.
By Lemma~\ref{lem:top-k-optimal-corners}\jessie{in the Appendix... remove ref for camera ready version}, we can see that for all $p \in \simplex$, there is some $u \in \{0,1\}^n$ so that $u \in \argmin_{u' \in \reals^d} \inprod{p}{L_2(u)}$.
With this, we show that $L_2$ actually embeds a discrete loss $\ell_2: \{0,1\}^n \to \reals^\Y_+$ defined on the corners of the hypercube below:
\begin{align}\label{eq:ell-2}
\ell_2(r,y) &= \Ind{y \not\in r} + \min\left(\frac{|r \setminus \{y\}|}{k}, 1\right)
\end{align}
We show that $L_2$ embeds $\ell_2$ in Lemma~\ref{lem:top-k-surrogate-embeds} \jessie{in the Appendix... remove ref for camera ready version}.
Moreover, by Proposition~\ref{prop:embed-bayes-risks}, we know $\risk{L_2} = \risk{\ell_2}$.

While $L_2$ given by~\citet{lapin2018analysis} is not consistent with respect to top-$k$ loss, we do show a discrete loss $\ell_2$ so that $L_2$ is consistent with respect to $\ell_2$.

Taking $k = 1$, we see that $\vec{0} \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ with $\max_y p_y \leq 1/2$, while $\risk{\ell_{top-1}}(p) = 1 - \max_y p_y < 1$.
First, we see that $\ell(\vec{0}, \bar{p}) = 1$ for all $p \in \simplex$.
Taking any basis vector $e_y$, the expected loss $\ell_2(e_y, p) = 2 (1-p_y) \geq 1$ by assumption on $p$. 
For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r,y) \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
Therefore, we have that $\vec{0} \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$.

@@

\section{Top-$k$}
Throughout this section, consider the surrogate and discrete loss $L_2(u,y)~:=~\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+$ and $\ell_2(r,y) = \Ind{y \not\in r} + \min\left(\frac{|r \setminus \{y\}|}{k}, 1\right)$ given in Equations~\ref{eq:L-2-surrogate} and~\ref{eq:ell-2}, respectively.

Observe that $L_2$ is a polyhedral loss, since it can be written as the pointwise max of $\binom{n}{k} +1$ terms, where the $\binom{n}{k}$ terms are selecting the $k$ elements of $u + \indopp(y)$ with highest weight.
By Lemma~\ref{lem:poly-loss-poly-risk}, we then know that $\risk{L_2}$ is also polyhedral.

\begin{lemma}\label{lem:top-k-optimal-corners}
	For all $p \in \simplex$, there is a $u \in \{0,1\}^n$ such that $u \in \argmin_{u'} \inprod{p}{L_2(u)}$.
\end{lemma}
\begin{proof}

\end{proof}

\begin{lemma}\label{lem:top-k-surrogate-embeds}
$L_2:\reals^n \to \reals^\Y_+$ embeds $\ell_2:2^{[n]} \to \reals^\Y_+$.
\end{lemma}
\begin{proof}
  We can verify $\ell_2(r) = L_2(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{0,1\}^n$.
  Moreover, by Lemma~\ref{lem:top-k-optimal-corners} and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L_2(u)} = \min_{u \in \{0,1\}^n} \inprod{p}{L_2(u)} = \min_{r \subseteq}\ell_2(r)$, and so we conclude that $L_2$ embeds $\ell_2$.
\end{proof}

>>>>

We have now seen the tight relationship between polyhedral losses and embeddings; in particular, every polyhedral loss embeds some discrete loss.
The embedding itself tells us how to map the embedded points back to the discrete reports (map $\varphi(r)$ to $r$), but it is not clear when this map can be extended to the remaining reports to obtain a full link function, and whether such a link can lead to consistency.

In this section, we give a construction to generate calibrated links for \emph{any} polyhedral loss.
The first step is to give a link $\psi$ such that minimizing expected surrogate loss, followed by applying $\psi$, minimizes expected original loss.

\begin{proposition}\label{prop:embed-link}
  Let $L$ elicit $\Gamma:\simplex\toto\reals^d$ which embeds a finite property $\gamma$.
  Then there exists a link function $\psi: \reals^d \to \R$ that is \emph{valid} in the sense that, for any $p$, if $u$ minimizes expected $L$-loss (i.e. $u \in \Gamma(p)$, then $\psi(u)$ minimizes expected $\ell$-loss (i.e. $\psi(u) \in \gamma(p)$).
\end{proposition}
\begin{proof}
  Let $\gamma:\simplex\toto\R$.
  Proposition~\ref{prop:embed-trim} gives us that $\trim(\Gamma) = \{\gamma_r : r\in\R\}$.
  \bo{Have we defined trim already? Or is this proof going to the appendix?}
  Then for any $u\in\reals^d$, there is some $r\in\R$ such that $\Gamma_u \subseteq \gamma_r$.
  The link $\psi : \reals^d \to \R$ which encodes these choices is valid.
\end{proof}

However, this guarantee is weak: for a given $p$, $u$ could be arbitrarily close to optimal with respect to $L$, yet $\psi(u)$ could remain arbitrarily wrong.
This kind of problem is ruled out by constructing a $\psi$ that is furthermore \emph{calibrated} (Definition \ref{def:calibrated}).
The main construction begins from the intuitive approach of ``thickening'' the embedding points, by causing all $u$ in the neighborhood of an embedding point to link back to the corresponding report.
However, we must consider cases where $\Gamma(p)$, the set of optimal $u$, is larger than just one embedding point.
These cases have a nice structure:

\begin{lemma}
  \label{lem:polyhedral-range-gamma}
  Let $L:\reals^d\to\reals_+^\Y$ be a polyhedral loss, and let $\Gamma = \prop{L}$.
  Then the range of $\Gamma$, $\U = \Gamma(\simplex) = \{\Gamma(p) \subseteq \reals^d | p\in\simplex\}$, is a finite set of closed polyhedra.
\end{lemma}
For each of these finitely many $U \in \U$, we let $R_U$ be the set of reports whose embedding points lie in $U$.
The idea will be to map points in $U$, as well as points nearby, to something in $R_U$
%We show in the full version \bo{okay wording?} that for any $p$, if $U = \Gamma(p)$, then $R_U = \gamma(p)$.

@@

\begin{proof}[Sketch]
  \emph{Well-defined:} Recall that Proposition \ref{prop:embed-link} shows that we can always consistently map each point of $U$ to some member of $R_U$.
  The key point was that if some collection such as $U,U',U''$ overlap at a $u$, then their report sets $R_U, R_{U'}, R_{U''}$ also overlap, so there is a valid choice $r = \psi(u)$.
  Now, we thicken all sets $U \in \U$ by a small enough $\epsilon$ and show that if the \emph{thickened} sets overlap at $u$, then $U,U',U''$ themselves overlap, so again $R_U, R_{U'}, R_{U''}$ overlap and there is a valid chioce $r = \psi(u)$.

  \emph{Calibrated:} By construction of the thickened link, if $u$ maps to a wrong report, i.e. $\psi(u) \not\in \gamma(p)$, then $u$ must have at least distance $\epsilon$ to the optimal set $U = \Gamma(p)$.
  We then show that the minimal gradient of the expected loss along any direction away from $U$ is lower-bounded, giving a constant excess expected loss at $u$.
\end{proof}

@@

  The top-$k$ classification problem takes as input a set $r \in \R := \{r \subseteq [n] : |r| \leq k\}$, and observes outcome $y \in [n]$.
  We define the top-$k$ loss $\ell_{top-k}:\R \to \reals$ by
\begin{align}\label{eq:top-k}
\ell_{top-k}(r,y) = \Ind{y \not \in r}
\end{align}

  \citet{lapin2016loss, lapin2018analysis}, provide two surrogate losses for measuring top-$k$ error, one of which is shown to be convex, but their consistency is left as an open question.
  \citet{yang2018consistency} show the inconsistency of their losses, and create a new surrogate loss that is consistent for the top-$k$ problem, but it should be noted that Yang and Koyejo's surrogate is not convex.
  They also show that Lapin's surrogates are not consistent, but without too much intuition why.
  Following the notation of Yang, we call Lapin's~\cite{lapin2018analysis} convex surrogate $L_2$, and focus on this.
  Here, we assume $r \subseteq [n]$ and $k < n$ is a constant.

 In order to do this, we want to show that for all $p \in \simplex$, $\inf_{u' \in \reals^n} \inprod{p}{L_2(u')} = \min_{u' \in \{0,1\}^n} \inprod{p}{L_2(u')}$. (Lemma~\ref{lem:top-k-optimal-corners}.) \jessiet{Appendix ref will be broken}
Then, we show there is a discrete loss $\ell_2 : \{0,1\}^n \to \reals^\Y_+$ that is embedded by $L_2$.  (Lemma~\ref{lem:top-k-surrogate-embeds}.)\jessiet{Appendix ref will be broken}

  We conclude with a concrete example: $k=1$.
  One can see that the $\ell_{top-1} = \ell_{0-1}$, and we characterize the distributions $p \in \simplex$ where $\risk{L_2}(p) = \risk{\ell_2}(p) \neq \risk{\ell_{top-k}}(p)$.

@@

We are given the surrogate 
\begin{align}\label{eq:L-2-surrogate}
L_2(u,y) := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+
\end{align}
where $x_{[i]}$ is the $i^{th}$ largest component of $x$.
By Lemma~\ref{lem:top-k-optimal-corners}\jessiet{in the Appendix... remove ref for camera ready version}, we can see that for all $p \in \simplex$, there is some $u \in \{0,1\}^n$ so that $u \in \argmin_{u' \in \reals^d} \inprod{p}{L_2(u)}$.
With this, we show that $L_2$ actually embeds the discrete loss $\ell_2: \R \to \reals^\Y_+$, where $\R := \{r \subseteq 2^{[n]} : |r| \leq k \}$, given below:
\begin{align}\label{eq:ell-2}
\ell_2(r,y) &= \Ind{y \not\in r} + \frac{|r \setminus \{y\}|}{k}
\end{align}
We show that $L_2$ embeds $\ell_2$ in Lemma~\ref{lem:top-k-surrogate-embeds} \jessiet{in the Appendix... remove ref for camera ready version}.
Moreover, by Proposition~\ref{prop:embed-bayes-risks}, we know $\risk{L_2} = \risk{\ell_2}$.

While $L_2$ given by~\citet{lapin2018analysis} is not consistent with respect to top-$k$ loss, we can at least show a discrete loss $\ell_2$ where $L_2$ is consistent with respect to $\ell_2$, applying Theorem~\ref{thm:embed-implies-calibrated-link}.\jessiet{TODO: Make ref}

Now to see a set of distributions where $\risk{L_2}(p) \neq \risk{\ell_{top-1}}(p)$, we claim that $\vec{0}~\in~\argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')} \implies \risk{\ell_2}(p) = 1$ for any $p$ with $\max_y p_y \leq 1/2$.
Additionally $\risk{\ell_{top-1}}(p) = 1 - \max_y p_y < 1$.

First, $\ell(\vec{0}, \bar{p}) = 1$ for all $p \in \simplex$.
Taking the set $\{y\}$ for any $y \in \Y$, the expected loss $\ell_2(\{y\}, p) = 2 (1-p_y) \geq 1$ by assumption on $p$. 
For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r,y) \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
Therefore, we have that $\vec{0} \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$.

@@

\section{Top-$k$ surrogate}
Throughout this section, consider the surrogate and discrete loss $L_2(u,y)~:=~\left(\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+$ and $\ell_2(r,y) = \Ind{y \not\in r} + \frac{|r \setminus \{y\}|}{k}$ given in Equations~\ref{eq:L-2-surrogate} and~\ref{eq:ell-2}, respectively.

Observe that $L_2$ is a polyhedral loss, since it can be written as the pointwise max of $\binom{n}{k} +1$ terms, where the $\binom{n}{k}$ terms are selecting the $k$ elements of $u + \ones - e_y$ with highest weight.
By Lemma~\ref{lem:poly-loss-poly-risk}, we then know that $\risk{L_2}$ is also polyhedral.

\begin{lemma}\label{lem:top-k-optimal-corners}
	For all $p \in \simplex$, there is a $u \in \{ u \in \{0,1\}^n : \|u\|_1 \leq k\}$ such that $u \in \argmin_{u' \in \reals^d} \inprod{p}{L_2(u)}$.
\end{lemma}
\begin{proof}[Sketch]
  First, observe that $L_2$ is invariant in the $\ones$ direction, so without loss of generality, we can restrict our consideration to reports so that $u_{[n]} = 0$.

  We first claim that for all $y \in \Y$, the minimum of $L(u,y)$ is achieved by some $u \in [0,1]^n$.
  Observe that if $u_{[1]} \geq u_{[n]} + 1$, the value of the loss gets ``cropped'' and will be the same by reporting $u_{[1]} = u_{[n]} + 1 = 1$. \jessiet{Flesh this out later.}
  Since $u_{[n]} = 0 \implies u_{[1]} \in [0,1]$, we can then observe there is some $u \in \argmin_{u' \in \reals^d} L_2(u',y)$ so that $u \in [0,1]^n$.

  Now, since we can restrict to the closed hypercube, we claim that we can rewrite the surrogate $L_2$ as 
  \begin{align*}
  	\inprod{p}{L_2(u)} &:= \frac{1}{k} \left( \sum_{i=1}^k (1 - p_{[i]}) u_{[i]} + u_{[k+1]} \sum_{i=1}^k p_{[i]} \right) + 1 - \inprod{p}{u}
  \end{align*}

  To verify this, observe the first term of the summation comes from the probability that you incorrectly did not have $u_y$ in the top-$k$, and the second term comes from the $(k+1)^{th}$ term being added when $u_y$ is correctly in the top $k$ (and therefore not an averaged term, since $u_y \in [0,1]$.) 
  The $1$ is pulled out from the average, and $\inprod{p}{u}$ comes from the $-u_y$ term form each $y$ by linearity of expectation.
  We don't need to concern ourselves with the max operator since $\frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \geq 0$ for all $y \in \Y$ on $[0,1]^n$.

  Since, for $u_{[j]}$ and $j \geq k+2$, the expected loss only decreases by increasing $u_{[j]}$ as much as possible, we increase them to $u_{[k+1]}$.
  By our invariance in the $\ones$ direction, we can re-shift our values if necessary, setting $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$.

  Substituting these values into the expected loss, we now observe a linear objective over the hypercube, which is known to be optimal on some corner of the hypercube, so the rest of $u_{[1]}, u_{[2]}, \ldots, u_{[k]} \in \{0,1\}$.
  Therefore, for all $p \in \simplex$, there is a report $u \in \{0,1\}^n$ optimizing $L_2$.
  Moreover, for this optimal $u$, observe $\|u\|_{1} \leq k$ since we have $u_{[k+1]} = u_{[k+2]} = \ldots = u_{[n]} = 0$.
\end{proof}

\begin{lemma}\label{lem:top-k-surrogate-embeds}
$L_2:\reals^n \to \reals^\Y_+$ embeds $\ell_2:2^{[n]} \to \reals^\Y_+$.
\end{lemma}
\begin{proof}
  We can verify $\ell_2(r) = L_2(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{ u \in \{0,1\}^n : \|u\|_1 \leq k \}$.
  Moreover, by Lemma~\ref{lem:top-k-optimal-corners} and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L_2(u)} = \min_{u \in \varphi(\R)} \inprod{p}{L_2(u)} = \min_{r \in \R}\inprod{p}{\ell_2(r)}$ for all $p \in \simplex$, and so we conclude that $L_2$ embeds $\ell_2$.
\end{proof}
